---
title: "Constant Mean f GAEMM Runs"
author: "Scott Yanco"
date: "9/24/2020"
output: pdf_document
---

# Initialization

Load required libraries
```{r messages = F}
#data tools
library(data.table)
library(tidyverse)

#plotting tools
library(ggridges)
library(wesanderson)
library(reshape2)
library(viridis)
library(cowplot)
library(ggthemes)
```

Required source code from GAEMM.  *IMPORTANT NOTE*: The paths in the chunk below are relative to my personal repository structure - the GAEMM repository must first be cloned from github (https://github.com/allipierce/GAEMM) and then the file paths below modified to wherever it's been downloaded.  Eventually GAEMM will be made into a package that cn be installed directly and loaded using the `library()` command.
```{r}
#source code for GAEMM 
source("../../GAEMM/R/expendenergy.R")
source("../../GAEMM/R/move.R")
source("../../GAEMM/R/reproduce.R")
source("../../GAEMM/R/randomizetraits.R")
source("../../GAEMM/R/GAEMM.R")
source("../../GAEMM/R/seasonalworld.R")
source("../../GAEMM/R/results_functions.R")

#source code for ggplot themes
source("../../ggplot_theme_Publication-master/ggplot_theme_Publication-master/R/ggplot_theme_Publication.R")
```


# Theoretical Model with constant annual mean resource availability across latitudes (each latitude still expriences varying amplitudes of seasonality, but all have mean f = 0.7)

## Model Runs

### High Seasonality: Loc=175

Set model parameters
```{r}
modelparams_175_cf <- list(worldsize = 181,
                    popsize = 10000,
                    steps = 365,
                    generations = 500,
                    g_max = 150,
                    l_p = 0.7,
                    kappa_min = 0.1,
                    kappa_max = 0.9999,
                    maxmoves = 45,
                    e_traits = c("g", "kappa", "m"),
                    startloc = 175,
                    replacerate = 0.8,
                    crossoverrate = 0.5,
                    mutationrate = 0.1,
                    selectiontype = "sus",
                    worldedge = "stop",
                    ncrosspoints = 2,
                    movement = F,
                    fillfirst = T)
```

Make the seasonal world - uses model params object, but world will be identical across treatments
```{r}
world <- make_seasonalworld(modelparams_175_cf, type = "sine", meanf = 0.7)
```

Run model
```{r}
totalresults_175_cf <- GAEMM(modelparams_175_cf, world, lastonly = T)
```

### Moderate Seasonality: Loc=133

Set model parameters
```{r}
modelparams_133_cf <- list(worldsize = 181,
                    popsize = 10000,
                    steps = 365,
                    generations = 500,
                    g_max = 150,
                    l_p = 0.7,
                    kappa_min = 0.1,
                    kappa_max = 0.9999,
                    maxmoves = 45,
                    e_traits = c("g", "kappa", "m"),
                    startloc = 133,
                    replacerate = 0.8,
                    crossoverrate = 0.5,
                    mutationrate = 0.1,
                    selectiontype = "sus",
                    worldedge = "stop",
                    ncrosspoints = 2,
                    movement = F,
                    fillfirst = T)
```

Run model
```{r}
totalresults_133_cf <- GAEMM(modelparams_133_cf, world, lastonly = T)
```

### No Seasonality: Loc=91

Set model parameters
```{r}
modelparams_91_cf <- list(worldsize = 181,
                    popsize = 10000,
                    steps = 365,
                    generations = 500,
                    g_max = 150,
                    l_p = 0.7,
                    kappa_min = 0.1,
                    kappa_max = 0.9999,
                    maxmoves = 45,
                    e_traits = c("g", "kappa", "m"),
                    startloc = 91,
                    replacerate = 0.8,
                    crossoverrate = 0.5,
                    mutationrate = 0.1,
                    selectiontype = "sus",
                    worldedge = "stop",
                    ncrosspoints = 2,
                    movement = F,
                    fillfirst = T)
```


Run model
```{r}
totalresults_91_cf <- GAEMM(modelparams_91_cf, world, lastonly = T)
```


## Process Data

First filter raw model output to include only "alive" solutions and those that have been evaluated in at least one generation, also remove entirely identical solutions

```{r}
filt_175_cf <- totalresults_175_cf %>%
  filter(alive == T & age > 1) %>% #remove dead and unevaluated solutions
  distinct() #remove complete duplicated solutions

filt_133_cf <- totalresults_133_cf %>%
  filter(alive == T & age > 1) %>% #remove dead and unevaluated solutions
  distinct() #remove complete duplicated solutions

filt_91_cf <- totalresults_91_cf %>%
  filter(alive == T & age > 1) %>% #remove dead and unevaluated solutions
  distinct() #remove complete duplicated solutions
```

  
Combine datasets into single df with start as a factor
```{r}
comb_cf <- rbind(filt_91_cf, filt_133_cf, filt_175_cf)
comb_cf$startloc <- as.factor(comb_cf$startloc)
```

Save out data files - note the relative filepath outside the git repository - may need to be adjusted depending on needs/personal preferences.

We save the filtered, combined object; unfiltered model results; parameterization objects; and, worlds.
```{r}
filename <- paste0("../data/data_nomove__cf",Sys.Date(),".rdata")
save(comb_cf, 
     modelparams_91_cf, modelparams_133_cf, modelparams_175_cf,
     totalresults_91_cf, totalresults_133_cf, totalresults_175_cf,
     world, 
     file = filename)
```

Make df of world for plotting and reduce to only locations used
```{r}
worlddf_cf <- melt(world, varnames = c("step", "location"))

#filter to only treatment locations
worlddf_cf_starts <- worlddf_cf %>% 
  filter(location==91|location==133|location==175)

#convert starts to factor variable and reorder
worlddf_cf_starts$location <- as.factor(worlddf_cf_starts$location)
worlddf_cf_starts$location <- fct_relevel(worlddf_cf_starts$location, "91", "133", "175")
```

#Plotting and Analysis

## Figure 1 - Experimental Design Heuristic

Making a plot of the seasonal fluctuations in resources over time for each location treatment.

First calculate annual mean resource levels for each location (except 91, since raw resources == mean resources).
```{r}
worldmeans_cf <- worlddf_cf_starts %>% 
  group_by(location) %>% 
  summarise(meanf = mean(value)) %>% 
  filter(location != "91")
```

Make a plot of experimental design wrt the world

```{r}
#set palette
pal_3 <- wes_palette("Darjeeling1", 3)

(worldplot <- ggplot() +
    geom_line(data = worlddf_cf_starts,
              aes(x=step, y = value, group=location, color = location), 
              size = 1.5) +
    geom_segment(data = worldmeans_cf,
                 aes(y = meanf, x = 0, yend = meanf, xend = 365, color = location),
                 size = 1, linetype = "twodash") +
    scale_color_manual(values = pal_3, name = "Model Latitude") +
    ylim(c(0,1)) +
    xlab("Time Step") +
    ylab("Available Resources") +
    theme_Publication() +
    theme(legend.position = "none"))
```
## Figure 2 - Model Results

Make 2 panel plot of model results for kappa (somatic allocation fraction) and g (energy investment ratio)

### Somatic allocation fraction (kappa) panel
```{r}
(plot_kappa_cf <- ggplot(comb_cf, aes(y=startloc, x = kappa, color = startloc)) +
   geom_jitter(alpha = 0.7) +
   # scale_fill_viridis(name = "Kappa") +
   scale_color_manual(values = pal_3) + #use same palette as worldplot
   scale_discrete_manual(aesthetics="point_color", values = pal_3) +
   # geom_smooth(method = "lm") +
   xlab(expression(paste("Somatic Allocation Fraction (", kappa, ")"))) +
   ylab("Latitude") +
   # scale_point_fill_hue(l = 70) +
   theme_Publication()+
   theme(legend.position = "none"))
```
### Energy investment ratio (g) panel
```{r}
(plot_g_cf <- ggplot(comb_cf, aes(y=startloc, x = g,color = startloc)) +
  geom_jitter(alpha = 0.7) +
  scale_color_manual(values = pal_3) +
  scale_discrete_manual(aesthetics="point_color", values = pal_3) +
  xlab(expression(paste("Energy Investment Ratio (", g, ")"))) +
  ylab("Latitude") +
  theme_Publication()+
  theme(legend.position = "none"))
```
### Combine into single panel

```{r}
(fig2_comb_cf <- plot_grid(plot_kappa_cf, plot_g_cf, labels = c("A", "B"), ncol = 1))
```