---
title: "seasonality_and_LH"
author: "Scott Yanco"
date: "9/23/2020"
output: pdf_document
---

# Initialization

Load required libraries
```{r messages = F}
#data tools
library(data.table)
library(tidyverse)
library(R.matlab)

#plotting tools
library(ggridges)
library(wesanderson)
library(reshape2)
library(viridis)
library(cowplot)
library(ggthemes)
```

Required source code from GAEMM.  *IMPORTANT NOTE*: The paths in the chunk below are relative to my personal repository structure - the GAEMM repository must first be cloned from github (https://github.com/allipierce/GAEMM) and then the file paths below modified to wherever it's been downloaded.  Eventually GAEMM will be made into a package that cn be installed directly and loaded using the `library()` command.
```{r}
#source code for GAEMM 
source("../../GAEMM/R/expendenergy.R")
source("../../GAEMM/R/move.R")
source("../../GAEMM/R/reproduce.R")
source("../../GAEMM/R/randomizetraits.R")
source("../../GAEMM/R/GAEMM.R")
source("../../GAEMM/R/seasonalworld.R")
source("../../GAEMM/R/results_functions.R")

#source code for ggplot themes
source("../../ggplot_theme_Publication-master/ggplot_theme_Publication-master/R/ggplot_theme_Publication.R")

#source code for AmP accessor functions
source("AmP_functions.R")
```


# Theoretical Model

## Model Runs

### High Seasonality: Loc=175

Set model parameters
```{r}
modelparams_175 <- list(worldsize = 181,
                    popsize = 10000,
                    steps = 365,
                    generations = 500,
                    g_max = 150,
                    l_p = 0.7,
                    kappa_min = 0.1,
                    kappa_max = 0.9999,
                    maxmoves = 45,
                    e_traits = c("g", "kappa", "m"),
                    startloc = 175,
                    replacerate = 0.8,
                    crossoverrate = 0.5,
                    mutationrate = 0.1,
                    selectiontype = "sus",
                    worldedge = "stop",
                    ncrosspoints = 2,
                    movement = F,
                    fillfirst = T)
```

Make the seasonal world - uses model params object, but world will be identical across treatments
```{r}
world_175 <- make_seasonalworld(modelparams_175, type = "sine", meanrange = c(0.7,0.85))
```

Run model
```{r}
totalresults_175 <- GAEMM(modelparams_175, world_175, lastonly = T)
```

### Moderate Seasonality: Loc=133

Set model parameters
```{r}
modelparams_133 <- list(worldsize = 181,
                    popsize = 10000,
                    steps = 365,
                    generations = 500,
                    g_max = 150,
                    l_p = 0.7,
                    kappa_min = 0.1,
                    kappa_max = 0.9999,
                    maxmoves = 45,
                    e_traits = c("g", "kappa", "m"),
                    startloc = 133,
                    replacerate = 0.8,
                    crossoverrate = 0.5,
                    mutationrate = 0.1,
                    selectiontype = "sus",
                    worldedge = "stop",
                    ncrosspoints = 2,
                    movement = F,
                    fillfirst = T)
```

Make the seasonal world - uses model params object, but world will be identical across treatments
```{r}
world_133 <- make_seasonalworld(modelparams_133, type = "sine", meanrange = c(0.7,0.85))
```

Run model
```{r}
totalresults_133 <- GAEMM(modelparams_133, world_133, lastonly = T)
```

### No Seasonality: Loc=91

Set model parameters
```{r}
modelparams_91 <- list(worldsize = 181,
                    popsize = 10000,
                    steps = 365,
                    generations = 500,
                    g_max = 150,
                    l_p = 0.7,
                    kappa_min = 0.1,
                    kappa_max = 0.9999,
                    maxmoves = 45,
                    e_traits = c("g", "kappa", "m"),
                    startloc = 91,
                    replacerate = 0.8,
                    crossoverrate = 0.5,
                    mutationrate = 0.1,
                    selectiontype = "sus",
                    worldedge = "stop",
                    ncrosspoints = 2,
                    movement = F,
                    fillfirst = T)
```

Make the seasonal world - uses model params object, but world will be identical across treatments
```{r}
world_91 <- make_seasonalworld(modelparams_91, type = "sine", meanrange = c(0.7,0.85))
```

Run model
```{r}
totalresults_91 <- GAEMM(modelparams_91, world_91, lastonly = T)
```


## Process Data

First filter raw model output to include only "alive" solutions and those that have been evaluated in at least one generation, also remove entirely identical solutions

```{r}
filt_175 <- totalresults_175 %>%
  filter(alive == T & age > 1) %>% #remove dead and unevaluated solutions
  distinct() #remove complete duplicated solutions

filt_133 <- totalresults_133 %>%
  filter(alive == T & age > 1) %>% #remove dead and unevaluated solutions
  distinct() #remove complete duplicated solutions

filt_91 <- totalresults_91 %>%
  filter(alive == T & age > 1) %>% #remove dead and unevaluated solutions
  distinct() #remove complete duplicated solutions
```

  
Combine datasets into single df with start as a factor
```{r}
comb <- rbind(filt_91, filt_133, filt_175)
comb$startloc <- as.factor(comb$startloc)
```

Save out data files - note the relative filepath outside the git repository - may need to be adjusted depending on needs/personal preferences.

We save the filtered, combined object; unfiltered model results; parameterization objects; and, worlds.
```{r}
filename <- paste0("../data/data_nomove_",Sys.Date(),".rdata")
save(comb, 
     modelparams_91, modelparams_133, modelparams_175,
     totalresults_91, totalresults_133, totalresults_175,
     world_91, world_133, world_175, 
     file = filename)
```

Make df of world for plotting and reduce to only locations used
```{r}
#uses world_175 but could use any since word params are identical across treatments
worlddf <- melt(world_175, varnames = c("step", "location"))

#filter to only treatment locations
worlddf_starts <- worlddf %>% 
  filter(location==91|location==133|location==175)

#convert starts to factor variable and reorder
worlddf_starts$location <- as.factor(worlddf_starts$location)
worlddf_starts$location <- fct_relevel(worlddf_starts$location, "91", "133", "175")
```

# Empirical Analysis

Load in the Add-my-Pet (AmP) database.
```{r}
deb <- readMat("../data/AmPdata/allStat.mat")
```

Extract desired variables into data.frame object.

```{r}
#pull species list over which to iterate accessor functions
splist <- unlist(labels(deb$allStat))
splist <- splist[1:(length(splist)-2)] #remove last two entries which are not species

#extract ecoCode vector
eco <- lapply(splist, getPar, par = "ecoCode", deb_data = deb)

deb_df <- data.frame(model = unlist(lapply(splist, getPar, par ="model", deb_data = deb)), #which version of DEB model
                      complete = unlist(lapply(splist, getPar, par ="COMPLETE", deb_data = deb)), #model completeness score                      
                      climate = unlist(lapply(eco, getPrimeClim)), #primary climate code
                      climate_full = unlist(lapply(eco, getClim)), #full climate code
                      class = unlist(lapply(splist, getPar, par ="class", deb_data = deb)), #taxonomic class
                      family = unlist(lapply(splist, getPar, par ="family", deb_data = deb)), #taxonomic family
                      phylum = unlist(lapply(splist, getPar, par ="phylum", deb_data = deb)), #taxonomic phylum
                      mig = unlist(lapply(eco, getMig)), #migratory status
                      max_l = unlist(lapply(splist, getPar, par ="L.m", deb_data = deb)), #max length
                      E_0 = unlist(lapply(splist, getPar, par ="E.0", deb_data = deb)), #embryo cost
                      z = unlist(lapply(splist, getPar, par ="z", deb_data = deb)), #zoom factor
                      kappa = unlist(lapply(splist, getPar, par = "kap", deb_data = deb)), #somatic allocation fraction (kappa)
                      k = unlist(lapply(splist, getPar, par ="k", deb_data = deb)), #maintenance ratio (k)
                      g = unlist(lapply(splist, getPar, par ="g", deb_data = deb)), #energy investment ratio (g)
                      v=unlist(lapply(splist, getPar, par ="v", deb_data = deb))) #energy conductance (v)
```

Filter data:  remove marine species, use chordata only, only the 'std' DEB model with completeness >=2.
```{r}
deb_df_filt <- deb_df %>%
  filter(climate != "M" & model == "std" & complete >= 2 & phylum == "Chordata")
```


#Plotting and Analysis

## Figure 1 - Experimental Design Heuristic

Making a plot of the seasonal fluctuations in resources over time for each location treatment.

First calculate annual mean resource levels for each location (except 91, since raw resources == mean resources).
```{r}
worldmeans <- worlddf_starts %>% 
  group_by(location) %>% 
  summarise(meanf = mean(value)) %>% 
  filter(location != "91")
```

Make a plot of experimental design wrt the world

```{r}
#set palette
pal_3 <- wes_palette("Darjeeling1", 3)

(worldplot <- ggplot() +
    geom_line(data = worlddf_starts,
              aes(x=step, y = value, group=location, color = location), 
              size = 1.5) +
    geom_segment(data = worldmeans,
                 aes(y = meanf, x = 0, yend = meanf, xend = 365, color = location),
                 size = 1, linetype = "twodash") +
    scale_color_manual(values = pal_3, name = "Model Latitude") +
    ylim(c(0,1)) +
    xlab("Time Step") +
    ylab("Available Resources") +
    theme_Publication() +
    theme(legend.position = "none"))
```

## Figure 2 - Model Results

Make 2 panel plot of model results for kappa (somatic allocation fraction) and g (energy investment ratio)

### Somatic allocation fraction (kappa) panel
```{r}
(plot_kappa <- ggplot(comb, aes(y=startloc, x = kappa, color = startloc)) +
   geom_jitter(alpha = 0.7) +
   # scale_fill_viridis(name = "Kappa") +
   scale_color_manual(values = pal_3) + #use same palette as worldplot
   scale_discrete_manual(aesthetics="point_color", values = pal_3) +
   # geom_smooth(method = "lm") +
   xlab(expression(paste("Somatic Allocation Fraction (", kappa, ")"))) +
   ylab("Latitude") +
   # scale_point_fill_hue(l = 70) +
   theme_Publication()+
   theme(legend.position = "none"))
```
### Energy investment ratio (g) panel
```{r}
(plot_g <- ggplot(comb, aes(y=startloc, x = g,color = startloc)) +
  geom_jitter(alpha = 0.7) +
  scale_color_manual(values = pal_3) +
  scale_discrete_manual(aesthetics="point_color", values = pal_3) +
  xlab(expression(paste("Energy Investment Ratio (", g, ")"))) +
  ylab("Latitude") +
  theme_Publication()+
  theme(legend.position = "none"))
```
### Combine into single panel

```{r}
(fig2_comb <- plot_grid(plot_kappa, plot_g, labels = c("A", "B"), ncol = 1))
```

## Figure 3 - Empirical patterns from AmP

Set 5 color palette.
```{r}
pal_5 <- wes_palette("Darjeeling1", 5)
```

Three panel figure including kappa, g, and v.

### Somatic Allocation fraction (kappa) panel

```{r}
(amp_kap <- ggplot(deb_df_filt, aes(y=climate, x = kappa, color = climate)) +
   geom_jitter(alpha = 0.5) +
   scale_color_manual(values = pal_5) +
   scale_y_discrete(labels = c("Tropical", "Dry", "Temperate", "Boreal", "Polar")) +
   xlab(expression(paste("Somatic Allocation Fraction (", kappa, ")"))) +
   ylab("") +
   theme_Publication() +
   theme(legend.position = "none"))
```

### Energy conductance (v) panel

```{r}
(amp_v <- ggplot(deb_df_filt, aes(y=climate, x = v, color = climate)) +
   geom_jitter(alpha = 0.5) +
   scale_color_manual(values = pal_5) +
   scale_y_discrete(labels = c("Tropical", "Dry", "Temperate", "Boreal", "Polar")) +
   xlab(expression(paste("Energy Conductance (", italic(dot(v)), ")"))) +
   ylab("") +
   theme_Publication() +
   theme(legend.position = "none"))
```

### Energy investment ratio (g) panel

```{r}
(amp_g <- ggplot(deb_df_filt, aes(y=climate, x = g, color = climate)) +
  geom_jitter(alpha = 0.5) +
  scale_color_manual(values = pal_5) +
  scale_y_discrete(labels = c("Tropical", "Dry", "Temperate", "Boreal", "Polar")) +
  xlab(expression(paste("Energy Investment Ratio (g)"))) +
  ylab("") +
  theme_Publication() +
  theme(legend.position = "none"))
```

### Energy investment ratio (g) by kappa panel

```{r}
clim_labs <- c("Tropical", "Dry", "Temperate", "Boreal", "Polar")
names(clim_labs) <- c("A", "B", "C", "D", "E")

deb_df_filt$climate_relev <- fct_relevel(deb_df_filt$climate, "E", "D", "C", "B", "A")

(amp_g.kap <- ggplot(deb_df_filt, aes(y=g, x = kappa, color = climate)) +
    geom_point(alpha = 0.5) +
    scale_color_manual(values = pal_5) +
    # scale_(labels = c("Tropical", "Dry", "Temperate", "Boreal", "Polar")) +
    xlab(expression(paste(kappa))) +
    ylab(expression(paste("Energy Investment Ratio (g)"))) +
    theme_Publication() +
    theme(legend.position = "none") +
    facet_wrap(~climate_relev, ncol = 1, labeller = labeller(climate_relev = clim_labs)))
```

###Combine
```{r}
fig3_left <- plot_grid(amp_kap, amp_g, amp_v, labels = "AUTO", ncol = 1)
(fig3_comb <- plot_grid(fig3_left, amp_g.kap, labels = c("", "D"), ncol = 2, rel_widths = c(2,1)))
```

# Save Plots

Save Figure 1
```{r}
#as pdf
ggsave("../figures/fig-1_seasons.pdf", worldplot, width = 7, height = 4)

#as png
ggsave("../figures/fig-1_seasons.png", worldplot, width = 7, height = 4)
```

Save Figure 2
```{r}
#as pdf
ggsave("../figures/fig-2_model.pdf", fig2_comb, width = 7, height = 6)

#as png
ggsave("../figures/fig-2_model.png", fig2_comb, width = 7, height = 6)
```

Save Figure 3
```{r}
#as pdf
ggsave("../figures/fig-3_AmP.pdf", fig3_comb, width = 7, height = 7)

#as png
ggsave("../figures/fig-3_AmP.png", fig3_comb, width = 7, height = 7)
```

