---
title: "seasonality_and_LH"
author: "Scott Yanco"
date: "9/23/2020"
output: pdf_document
---

# Initialization

Load required libraries
```{r messages = F}
#data tools
library(data.table)
library(tidyverse)
library(R.matlab)

#plotting tools
library(ggridges)
library(wesanderson)
library(reshape2)
library(viridis)
library(cowplot)
library(ggthemes)

```

Required source code from GAEMM.  *IMPORTANT NOTE*: The paths in the chunk below are relative to my personal repository structure - the GAEMM repository must first be cloned from github (https://github.com/allipierce/GAEMM) and then the file paths below modified to wherever it's been downloaded.  Eventually GAEMM will be made into a package that cn be installed directly and loaded using the `library()` command.
```{r}
# #source code for GAEMM 
# source("../../GAEMM/R/expendenergy.R")
# source("../../GAEMM/R/move.R")
# source("../../GAEMM/R/reproduce.R")
# source("../../GAEMM/R/randomizetraits.R")
# source("../../GAEMM/R/GAEMM.R")
source("../../GAEMM/R/seasonalworld.R")
# source("../../GAEMM/R/results_functions.R")

#source code for analytic version of DEB model
source("debsim.R")

#source code for ggplot themes
source("../../ggplot_theme_Publication-master/ggplot_theme_Publication-master/R/ggplot_theme_Publication.R")

#source code for AmP accessor functions
source("AmP_functions.R")
```

Functions
```{r}
plotworld <- function(world){
  worlddata <- as.data.frame(world)
  names(worlddata) <- 1:181
  worlddata$t <- 1:365
  
  pivot_longer(worlddata, cols = -t, names_to = "loc", values_to = "f") %>%
    filter(loc == 175 | loc == 133 | loc == 91) %>%
    mutate(loc = factor(loc, levels = c("175", "133", "91"))) %>%
    ggplot() +
    geom_path(aes(x = t, y = f, color = loc, group = loc)) +
    scale_color_viridis_d(option = "C", end = 0.6, begin = 0.15) +
    theme_classic()
}

debsim <- function(modelparams, klow = 0.01, khigh = 0.999, glow = 0.01, ghigh = 5){
  
  kappa <- seq(klow, khigh, by = 0.01)
  g <- seq(glow, ghigh, by = 0.01)
  #EgEmrat <- seq(ratlow, rathigh, by = 0.01)
  
  #generate combinations of all kappas and gs
  #kappa_combo <- expand.grid(kappa = kappa, EgEmrat = EgEmrat)
  kappa_combo <- expand.grid(kappa = kappa, g = g)
  #kappa_combo$g <- kappa_combo$EgEmrat / kappa_combo$kappa
  
  #create initial data.frame    
  debdata <- data.frame(kappa = kappa_combo$kappa,
                        g = kappa_combo$g,
                        reserves = world[1, modelparams$startloc],
                        repro_reserves = 0,
                        l = modelparams$l_p,
                        l_p = modelparams$l_p,
                        startloc = modelparams$startloc,
                        loc = modelparams$startloc,
                        alive = T)
  
  #run steps for each parameter set
  for (i in 1:modelparams$steps) {
    debdata <- rundeb(debdata, i, world)
  }
  
  #return dataframe of values
  return(debdata)
}
```


# Theoretical Model

## Model Runs

### High Seasonality: Loc=175

Set model parameters
```{r}
modelparams_175 <- list(worldsize = 181,
                        steps = 365,
                        startloc = 175,
                        l_p = 0.7)
```

Make the seasonal world - uses model params object, but world will be identical across treatments
```{r}
world <- make_seasonalworld(modelparams_175, type = "sine", meanrange = c(0.7,0.85))
```

Run model
```{r}
totalresults_175 <- debsim(modelparams_175)
```

### Moderate Seasonality: Loc=133

Set model parameters
```{r}
modelparams_133 <- list(worldsize = 181,
                        steps = 365,
                        startloc = 133,
                        l_p = 0.7)
```

Run model
```{r}
totalresults_133 <- debsim(modelparams_133)
```

### No Seasonality: Loc=91

Set model parameters
```{r}
modelparams_91 <- list(worldsize = 181,
                       steps = 365,
                       startloc = 91,
                       l_p = 0.7)
```


Run model
```{r}
totalresults_91 <- debsim(modelparams_91)
```


## Process Data

Combine datasets into single df with start as a factor
```{r}
comb <- bind_rows(totalresults_175, totalresults_133, totalresults_91)
comb$startloc <- as.factor(comb$startloc)
comb$startloc <- fct_relevel(comb$startloc, "175", "133", "91")
```

First filter raw model output to include only "alive" solutions and those that have been evaluated in at least one generation, also remove entirely identical solutions

```{r}
comb <- comb %>%
  filter(alive == T) %>% #remove dead 
  distinct() #remove complete duplicated solutions
```

Save out data files - note the relative filepath outside the git repository - may need to be adjusted depending on needs/personal preferences.

We save the filtered, combined object; unfiltered model results; parameterization objects; and, worlds.
```{r}
filename <- paste0("../data/data_nomove_analytic",Sys.Date(),".rdata")
save(comb, 
     modelparams_91, modelparams_133, modelparams_175,
     totalresults_91, totalresults_133, totalresults_175,
     world, 
     file = filename)
```

Make df of world for plotting and reduce to only locations used
```{r}
worlddf <- melt(world, varnames = c("step", "location"))

#filter to only treatment locations
worlddf_starts <- worlddf %>% 
  filter(location==91|location==133|location==175)

#convert starts to factor variable and reorder
worlddf_starts$location <- as.factor(worlddf_starts$location)
worlddf_starts$location <- fct_relevel(worlddf_starts$location, "91", "133", "175")
```

# Empirical Analysis

Load in the Add-my-Pet (AmP) database.
```{r}
deb <- readMat("../data/AmPdata/allStat.mat")
```

Extract desired variables into data.frame object.

```{r}
#pull species list over which to iterate accessor functions
splist <- unlist(labels(deb$allStat))
splist <- splist[1:(length(splist)-2)] #remove last two entries which are not species

#extract ecoCode vector
eco <- lapply(splist, getPar, par = "ecoCode", deb_data = deb)

deb_df <- data.frame(model = unlist(lapply(splist, getPar, par ="model", deb_data = deb)), #which version of DEB model
                     complete = unlist(lapply(splist, getPar, par ="COMPLETE", deb_data = deb)), #model completeness score                      
                     climate = unlist(lapply(eco, getPrimeClim)), #primary climate code
                     climate_full = unlist(lapply(eco, getClim)), #full climate code
                     class = unlist(lapply(splist, getPar, par ="class", deb_data = deb)), #taxonomic class
                     family = unlist(lapply(splist, getPar, par ="family", deb_data = deb)), #taxonomic family
                     phylum = unlist(lapply(splist, getPar, par ="phylum", deb_data = deb)), #taxonomic phylum
                     mig = unlist(lapply(eco, getMig)), #migratory status
                     max_l = unlist(lapply(splist, getPar, par ="L.m", deb_data = deb)), #max length
                     Em = unlist(lapply(splist, getPar, par ="E.m", deb_data = deb)), #max reserve
                     pM = unlist(lapply(splist, getPar, par ="p.M", deb_data = deb)), #somatic maint
                     z = unlist(lapply(splist, getPar, par ="z", deb_data = deb)), #zoom factor
                     kappa = unlist(lapply(splist, getPar, par = "kap", deb_data = deb)), #somatic allocation fraction (kappa)
                     k = unlist(lapply(splist, getPar, par ="k", deb_data = deb)), #maintenance ratio (k)
                     g = unlist(lapply(splist, getPar, par ="g", deb_data = deb)), #energy investment ratio (g)
                     v=unlist(lapply(splist, getPar, par ="v", deb_data = deb))) #energy conductance (v)
```

Filter data:  remove marine species, use chordata only, only the 'std' DEB model with completeness >=2.
```{r}
deb_df_filt <- deb_df %>%
  filter(climate != "M" & complete >= 2) %>% 
  filter(class == "Mammalia" | class == "Aves" | class == "Reptilia")
```


#Plotting and Analysis

## Figure 1 - Experimental Design Heuristic

Making a plot of the seasonal fluctuations in resources over time for each location treatment.

First calculate annual mean resource levels for each location (except 91, since raw resources == mean resources).
```{r}
worldmeans <- worlddf_starts %>% 
  group_by(location) %>% 
  summarise(meanf = mean(value)) %>% 
  filter(location != "91")
```

Make a plot of experimental design wrt the world

Set "latitude" labels
```{r}
lat_labs <- c("High Seasonality", "Moderate Seasonality", "Aseasonal")
names(lat_labs) <- c("175", "133", "91")
```

```{r}
#set palette
pal_3 <- wes_palette("Darjeeling1", 3)

(worldplot <- ggplot() +
    geom_line(data = worlddf_starts,
              aes(x=step, y = value, group=location, color = location), 
              size = 1.5) +
    geom_segment(data = worldmeans,
                 aes(y = meanf, x = 0, yend = meanf, xend = 365, color = location),
                 size = 1, linetype = "twodash") +
    scale_color_manual(values = pal_3, name = "Model Latitude") +
    ylim(c(0,1)) +
    xlab("Time Step") +
    ylab("Available Resources") +
    theme_Publication() +
    theme(legend.position = "none"))
```

## Figure 2 - Model Results

```{r}
(kap_rp <- ggplot(comb) + geom_line(aes(x = kappa, y = repro_reserves, color = g, group = g), alpha = 0.5) +
   scale_color_viridis_c(name = expression(atop(atop("Energy", "Investment"), atop("Ratio (g)", "",))),
                         option = "inferno", direction = -1, end = 0.8) +
   
   facet_wrap(~startloc, ncol = 1, labeller = labeller(startloc = lat_labs)) +
   xlim(c(0,1)) +
   xlab(expression(paste("Somatic Allocation Fraction (", kappa, ")"))) +
   ylab("Reproductive Potential") +
   theme_few())
```


## Figure 3 - Empirical patterns from AmP

Set 5 color palette.
```{r}
pal_5 <- wes_palette("Darjeeling1", 5)
```


Three panel figure including kappa, g, and v.

### Somatic Allocation fraction (kappa) panel

Plot of kappa/Em by climate ("scaled POL")

```{r}
(amp_kap_E<- ggplot(deb_df_filt, aes(y=climate, x = (pM/Em), color = climate)) +
   geom_jitter(alpha = 0.5) +
   scale_color_manual(values = pal_5) +
   scale_y_discrete(labels = c("Tropical", "Dry", "Temperate", "Boreal", "Polar")) +
   xlab(expression(kappa[scaled])) +
   ylab("Climate") +
   theme_Publication() +
   theme(legend.position = "none"))
```

```{r}
# (amp_kap <- ggplot(deb_df_filt, aes(y=climate, x = kappa, color = climate)) +
#    geom_jitter(alpha = 0.5) +
#    scale_color_manual(values = pal_5) +
#    scale_y_discrete(labels = c("Tropical", "Dry", "Temperate", "Boreal", "Polar")) +
#    xlab(expression(paste("Somatic Allocation Fraction (", kappa, ")"))) +
#    ylab("") +
#    theme_Publication() +
#    theme(legend.position = "none"))
```

### Energy conductance (v) panel

```{r}
# (amp_v <- ggplot(deb_df_filt, aes(y=climate, x = v, color = climate)) +
#    geom_jitter(alpha = 0.5) +
#    scale_color_manual(values = pal_5) +
#    scale_y_discrete(labels = c("Tropical", "Dry", "Temperate", "Boreal", "Polar")) +
#    # xlab(expression(paste("Energy Conductance (", italic(dot(v)), ")"))) +
#    ylab("") +
#    theme_Publication() +
#    theme(legend.position = "none"))
```

### Energy investment ratio (g) panel

```{r}
(amp_g <- ggplot(deb_df_filt, aes(y=climate, x = g, color = climate)) +
   geom_jitter(alpha = 0.5) +
   scale_color_manual(values = pal_5) +
   scale_y_discrete(labels = c("Tropical", "Dry", "Temperate", "Boreal", "Polar")) +
   xlab(expression(paste("Energy Investment Ratio (g)"))) +
   ylab("") +
   theme_Publication() +
   theme(legend.position = "none"))
```

### Energy conductance (v) by kappa panel

```{r}
# clim_labs <- c("Tropical", "Dry", "Temperate", "Boreal", "Polar")
# names(clim_labs) <- c("A", "B", "C", "D", "E")
# 
# deb_df_filt$climate_relev <- fct_relevel(deb_df_filt$climate, "E", "D", "C", "B", "A")
# 
# (amp_g.kap <- ggplot(deb_df_filt, aes(y=v, x = kappa, color = climate)) +
#     geom_point(alpha = 0.5) +
#     scale_color_manual(values = pal_5) +
#     # scale_(labels = c("Tropical", "Dry", "Temperate", "Boreal", "Polar")) +
#     xlab(expression(paste(kappa))) +
#     ylab(expression(paste("Energy Conductance (", italic(dot(v)), ")"))) +
#     theme_Publication() +
#     theme(legend.position = "none") +
#     facet_wrap(~climate_relev, ncol = 1, labeller = labeller(climate_relev = clim_labs)))
```

###Combine
```{r}
(fig3 <- plot_grid(amp_kap_E, amp_g, labels = "AUTO", ncol = 1))
# (fig3_comb <- plot_grid(fig3_left, amp_g.kap, labels = c("", "D"), ncol = 2, rel_widths = c(2,1)))
```

##Figure 4 - Kappa*v by climate


# Save Plots

Save Figure 1
```{r}
#as pdf
ggsave("../figures/fig-1_seasons.pdf", worldplot, width = 7, height = 4)

#as png
ggsave("../figures/fig-1_seasons.png", worldplot, width = 7, height = 4)
```

Save Figure 2
```{r}
#as pdf
ggsave("../figures/fig-2_model.pdf", kap_rp, width = 7, height = 6)

#as png
ggsave("../figures/fig-2_model.png", kap_rp, width = 7, height = 6)
```

Save Figure 3
```{r}
#as pdf
ggsave("../figures/fig-3_AmP.pdf", fig3_left, width = 7, height = 7)

#as png
ggsave("../figures/fig-3_AmP.png", fig3_left, width = 7, height = 7)
```

Save Figure 4
```{r}
# #as pdf
# ggsave("../figures/fig-4_AmP_kap-v.pdf", amp_kap_v, width = 5, height = 5)
# 
# #as png
# ggsave("../figures/fig-4_AmP_kap-v.png", amp_kap_v, width = 5, height = 5)
```
# Descriptive Stats

## World Means

```{r}
worlddf_starts %>% 
  group_by(location) %>% 
  summarise(meanf = mean(value), min = min(value), max = max(value))
```


## Max kappa by lat

```{r}
print(paste0("Start = 91: ", max(comb$kappa[comb$startloc == "91"])))
print(paste0("Start = 133: ", max(comb$kappa[comb$startloc == "133"])))
print(paste0("Start = 175: ", max(comb$kappa[comb$startloc == "175"])))
```
## Min kappa by lat

```{r}
print(paste0("Start = 91: ", min(comb$kappa[comb$startloc == "91"])))
print(paste0("Start = 133: ", min(comb$kappa[comb$startloc == "133"])))
print(paste0("Start = 175: ", min(comb$kappa[comb$startloc == "175"])))
```

## Kappa at max reproductive potential
```{r}

comb %>% 
  group_by(startloc) %>% 
  filter(repro_reserves == max(repro_reserves)) %>% 
  select(startloc, kappa, g, repro_reserves)
```

# Constant mean f runs

## Model Runs

### High Seasonality: Loc=175

Set model parameters
```{r}
modelparams_175_cf <- list(worldsize = 181,
                           steps = 365,
                           startloc = 175,
                           l_p = 0.7)
```

Make the seasonal world - uses model params object, but world will be identical across treatments
```{r}
world <- make_seasonalworld(modelparams_175_cf, type = "sine", meanf = 0.75)
```

Run model
```{r}
totalresults_175_cf <- debsim(modelparams_175_cf)
```

### Moderate Seasonality: Loc=133

Set model parameters
```{r}
modelparams_133_cf <- list(worldsize = 181,
                           steps = 365,
                           startloc = 133,
                           l_p = 0.7)
```

Run model
```{r}
totalresults_133_cf <- debsim(modelparams_133_cf)
```

### No Seasonality: Loc=91

Set model parameters
```{r}
modelparams_91_cf <- list(worldsize = 181,
                          steps = 365,
                          startloc = 91,
                          l_p = 0.7)
```


Run model
```{r}
totalresults_91_cf <- debsim(modelparams_91_cf)
```

Combine datasets into single df with start as a factor
```{r}
comb_cf <- bind_rows(totalresults_175_cf, totalresults_133_cf,
                     totalresults_91_cf)
comb_cf$startloc <- as.factor(comb_cf$startloc)
comb_cf$startloc <- fct_relevel(comb_cf$startloc, "175", "133", "91")
```

First filter raw model output to include only "alive" solutions and those that have been evaluated in at least one generation, also remove entirely identical solutions

```{r}
comb_cf <- comb_cf %>%
  filter(alive == T) %>% #remove dead 
  distinct() #remove complete duplicated solutions
```

Save out data files - note the relative filepath outside the git repository - may need to be adjusted depending on needs/personal preferences.

We save the filtered, combined object; unfiltered model results; parameterization objects; and, worlds.
```{r}
filename <- paste0("../data/data_nomove_analytic_cf",Sys.Date(),".rdata")
save(comb, 
     modelparams_91_cf, modelparams_133_cf, modelparams_175_cf,
     totalresults_91_cf, totalresults_133_cf, totalresults_175_cf,
     world, 
     file = filename)
```

Make df of world for plotting and reduce to only locations used
```{r}
worlddf <- melt(world, varnames = c("step", "location"))

#filter to only treatment locations
worlddf_starts <- worlddf %>% 
  filter(location==91|location==133|location==175)

#convert starts to factor variable and reorder
worlddf_starts$location <- as.factor(worlddf_starts$location)
worlddf_starts$location <- fct_relevel(worlddf_starts$location, "91", "133", "175")
```

## Figure 1 - Experimental Design Heuristic

Making a plot of the seasonal fluctuations in resources over time for each location treatment.

First calculate annual mean resource levels for each location (except 91, since raw resources == mean resources).
```{r}
worldmeans <- worlddf_starts %>% 
  group_by(location) %>% 
  summarise(meanf = mean(value)) 


worldmeans$location <- fct_relevel(worldmeans$location, "175", "133", "91")
```

Make a plot of experimental design wrt the world

```{r}
#set palette
pal_3 <- wes_palette("Darjeeling1", 3)

(worldplot <- ggplot() +
    geom_line(data = worlddf_starts,
              aes(x=step, y = value, group=location, color = location), 
              size = 1.5) +
    scale_color_manual(values = pal_3, name = "Model Latitude") +
    ylim(c(0,1)) +
    xlab("Time Step") +
    ylab("Available Resources") +
    theme_Publication() +
    theme(legend.position = "none"))

(meanfs <- ggplot()+
    geom_segment(data = worldmeans,
                 aes(y = meanf, x = 0, yend = meanf, xend = 365, 
                     color = location),
                 size = 1, linetype = "twodash") +
    ylim(c(0,1)) +
    xlab("Time Step") +
    ylab("Mean Available Resources") +
    facet_wrap(~location, ncol = 1, 
               labeller = labeller(location = lat_labs)) +
    theme_Publication() +
    theme(legend.position = "none"))
```

Combine panels:

```{r}
(fig_cf_comb <- plot_grid(worldplot, meanfs, labels = c("A", "B"), ncol = 2, rel_widths = c(2,1.6)))
```


## Figure 2 - Model Results

```{r}
(kap_rp_cf <- ggplot(comb_cf) + 
   geom_line(aes(x = kappa, y = repro_reserves, color = g, group = g), 
             alpha = 0.5) +
   scale_color_viridis_c(name = expression(atop(atop("Energy", "Investment"),
                                                atop("Ratio (g)", "",))),
                         option = "inferno", direction = -1, end = 0.8) +
   facet_wrap(~startloc, ncol = 1, labeller = labeller(startloc = lat_labs)) +
   xlim(c(0,1)) +
   xlab(expression(paste("Somatic Allocation Fraction (", kappa, ")"))) +
   ylab("Reproductive Potential") +
   theme_few())
```

# Save Plots

Save Figure E1
```{r}
#as pdf
ggsave("../figures/fig-E1_seasons.pdf", fig_cf_comb, width = 7, height = 4)

#as png
ggsave("../figures/fig-E1_seasons.png", fig_cf_comb, width = 7, height = 4)
```

Save Figure E2
```{r}
#as pdf
ggsave("../figures/fig-E_model.pdf", kap_rp_cf, width = 7, height = 6)

#as png
ggsave("../figures/fig-E2_model.png", kap_rp_cf, width = 7, height = 6)
```
