---
title: "seasonality_and_LH"
author: "Scott Yanco"
date: "9/23/2020"
output: pdf_document
---

# Initialization

Load required libraries
```{r}
library(data.table)
library(tidyverse)
library(ggridges)
library(wesanderson)
library(reshape2)
library(viridis)
```

Required source code from GAEMM
```{r}
source("R/expendenergy.R")
source("R/move.R")
source("R/reproduce.R")
source("R/randomizetraits.R")
source("R/GAEMM.R")
source("R/seasonalworld.R")
source("R/results_functions.R")
```


# Model Runs

## High Seasonality: Loc=175

Set model parameters
```{r}
modelparams_175 <- list(worldsize = 181,
                    popsize = 10000,
                    steps = 365,
                    generations = 500,
                    g_max = 150,
                    l_p = 0.7,
                    kappa_min = 0.1,
                    kappa_max = 0.9999,
                    maxmoves = 45,
                    e_traits = c("g", "kappa", "m"),
                    startloc = 175,
                    replacerate = 0.8,
                    crossoverrate = 0.5,
                    mutationrate = 0.1,
                    selectiontype = "sus",
                    worldedge = "stop",
                    ncrosspoints = 2,
                    movement = F,
                    fillfirst = T)
```

Make the seasonal world - uses model params object, but world will be identical across treatments
```{r}
world_175 <- make_seasonalworld(modelparams_175, type = "sine", meanrange = c(0.7,0.85))
```

Run model
```{r}
totalresults_175 <- GAEMM(modelparams_175, world_175, lastonly = T)
```

## Moderate Seasonality: Loc=133

Set model parameters
```{r}
modelparams_175 <- list(worldsize = 181,
                    popsize = 10000,
                    steps = 365,
                    generations = 500,
                    g_max = 150,
                    l_p = 0.7,
                    kappa_min = 0.1,
                    kappa_max = 0.9999,
                    maxmoves = 45,
                    e_traits = c("g", "kappa", "m"),
                    startloc = 133,
                    replacerate = 0.8,
                    crossoverrate = 0.5,
                    mutationrate = 0.1,
                    selectiontype = "sus",
                    worldedge = "stop",
                    ncrosspoints = 2,
                    movement = F,
                    fillfirst = T)
```

Make the seasonal world - uses model params object, but world will be identical across treatments
```{r}
world_133 <- make_seasonalworld(modelparams_133, type = "sine", meanrange = c(0.7,0.85))
```

Run model
```{r}
totalresults_133 <- GAEMM(modelparams_133, world_133, lastonly = T)
```

## No Seasonality: Loc=91

Set model parameters
```{r}
modelparams_175 <- list(worldsize = 181,
                    popsize = 10000,
                    steps = 365,
                    generations = 500,
                    g_max = 150,
                    l_p = 0.7,
                    kappa_min = 0.1,
                    kappa_max = 0.9999,
                    maxmoves = 45,
                    e_traits = c("g", "kappa", "m"),
                    startloc = 91,
                    replacerate = 0.8,
                    crossoverrate = 0.5,
                    mutationrate = 0.1,
                    selectiontype = "sus",
                    worldedge = "stop",
                    ncrosspoints = 2,
                    movement = F,
                    fillfirst = T)
```

Make the seasonal world - uses model params object, but world will be identical across treatments
```{r}
world_91 <- make_seasonalworld(modelparams_91, type = "sine", meanrange = c(0.7,0.85))
```

Run model
```{r}
totalresults_91 <- GAEMM(modelparams_91, world_91, lastonly = T)
```


#Process Data

First filter raw model output to include only "alive" solutions and those that have been evaluated in at least one generation, also remove entirely identical solutions

```{r}
filt_175 <- totalresults_175 %>%
  filter(alive == T & age > 1) %>% #remove dead and unevaluated solutions
  distinct() #remove complete duplicated solutions

filt_133 <- totalresults_133 %>%
  filter(alive == T & age > 1) %>% #remove dead and unevaluated solutions
  distinct() #remove complete duplicated solutions

filt_91 <- totalresults_91 %>%
  filter(alive == T & age > 1) %>% #remove dead and unevaluated solutions
  distinct() #remove complete duplicated solutions
```

  
Combine datasets into single df with start as a factor
```{r}
comb <- rbind(filt_91, filt_133, filt_175)
comb$startloc <- as.factor(comb$startloc)
```

Save out data file
```{r}
filename <- paste0("data_nomove_",Sys.Date(),".rdata")
save(comb, file = filename)
```

Make df of world for plotting and dreduce to only locations used
```{r}
#uses world_175 but could use any since word params are identical across treatments
worlddf <- melt(world_175, varnames = c("step", "location"))

#filter to only treatment locations
worlddf_starts <- worlddf %>% 
  filter(location==91|location==133|location==175)

#convert starts to factor variable and reorder
worlddf_starts$location <- as.factor(worlddf_starts$location)
worlddf_starts$location <- fct_relevel(worlddf_starts$location, "91", "133", "175")
```

#Plotting and Analysis

## Figure 1 - Experiemntal Design Heuristic

Making a plot of the seasonal fluctuations in resources over time for each location treatment.

First calculate annual mean resource levels for each location (except 91, since raw resources == mean resources).
```{r}
worldmeans <- worlddf_starts %>% 
  group_by(location) %>% 
  summarise(meanf = mean(value)) %>% 
  filter(location != "91")
```

Make a plot of experimental design wrt the world

```{r}
#set palette
pal1 <- wes_palette("Darjeeling1", 3)

(worldplot <- ggplot() +
  geom_line(data = worlddf_starts,
            aes(x=step, y = value, group=location, color = location), 
            size = 2) +
  geom_segment(data = worldmeans,
             aes(y = meanf, x = 0, yend = meanf, xend = 365, color = location),
             size = 1.5, linetype = "twodash") +
  scale_color_manual(values = pal1, name = "Model Latitude") +
  xlab("Time Step") +
  ylab("Available Resources") +
  theme_minimal())
```

# Save Plots

Save Figure 1
